### 2019-03-15 notebooks/general_info.Rmd
added function to display data frames in notebook, and functions to get fully monomorphic things (i.e., monomorphic in class, attribue, and type). the counts are available in the notebook, just not being printed yet (you might need to open it with RStudio to see)

---
title: "Overview"
output: html_notebook
---

Ok.
First, load the full signatures, and make sure we source the data processing file.
There's an error here, not sure what's causing it.
This is a work-in-progress.

```{r}
source("../code/data_processing.R") # get the functions we need
```

```{r, echo=FALSE}
los <- list.files("../data/full_sigs", full.names=TRUE)
names <- list.files("../data/full_sigs")
names(los) <- substr(names, 1, nchar(names)-4)
los <- lapply(los, readRDS)
los <- remove_empties(los)
# annoying strange error that I should fix pops up
```

These are the updated counts.
```{r}
total_arguments <- apply_count_to_lopkg_c(los, args_count_all)
total_mono_type <- apply_count_to_lopkg_c(los, args_count_all_mono_type)
total_mono_attr <- apply_count_to_lopkg_c(los, args_count_all_mono_attr_pattern)
total_mono_clas <- apply_count_to_lopkg_c(los, args_count_all_mono_class)

total_functions <- apply_count_to_lopkg_c(los, length)
total_mono_type_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_type)
total_mono_attr_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_attr_pattern)
total_mono_clas_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_class)
```

This is how you get the top sigs.
```{r}
poly_type_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_type_args))
poly_type_args <- poly_type_args[order(-poly_type_args$count),] # sort by count
poly_type_args$perc <- round(poly_type_args$count/sum(poly_type_args$count) * 100, 2)
```

Then, pick out the ones you want.
The `format_df_for_print` function is just for letting the notebook display the data frame properly.
The type, attr, and class rows are lists of lists.
```{r}
format_df_for_print(poly_type_args[1:20,])
```

Similarly for attributes, and classes.
```{r}
poly_attr_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_attr_args))
poly_attr_args <- poly_attr_args[order(-poly_attr_args$count),] # sort by count
poly_attr_args$perc <- round(poly_attr_args$count/sum(poly_attr_args$count) * 100, 2)
```

```{r}
poly_class_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_class_args))
poly_class_args <- poly_class_args[order(-poly_class_args$count),] # sort by count
poly_class_args$perc <- round(poly_class_args$count/sum(poly_class_args$count) * 100, 2)
```

Printing is as before.
Note that ATM I also have attribute type information in here.
I'll get it to display just the names as well.
I figured this might be interesting, and I don't expect many functions to be polymorphic in this way.
```{r}
format_df_for_print(poly_attr_args[1:20,])
```

```{r}
format_df_for_print(poly_class_args[1:20,])
```