### 2019-03-23 notebooks/paper_results_section.Rmd
plots and new data, with vector/scalar dealt with

---
title: "Paper Results Section"
output: html_notebook
---

This Notebook walks through the results section of the paper.

First, administrivia.
```{r}
source("../code/data_processing.R") # get the functions we need

library(xtable)
```

Load the data.
```{r, echo=FALSE}
los <- list.files("../data/full_sigs", full.names=TRUE)
names <- list.files("../data/full_sigs")
names(los) <- substr(names, 1, nchar(names)-4)
los <- lapply(los, readRDS)
los <- remove_empties(los)
# annoying strange error that I should fix pops up
```

Get the counts.
```{r}
# argument counts
total_arguments <- apply_count_to_lopkg_c(los, args_count_all)
total_mono_type <- apply_count_to_lopkg_c(los, args_count_all_mono_type)
total_mono_attr <- apply_count_to_lopkg_c(los, args_count_all_mono_attr_pattern)
total_mono_clas <- apply_count_to_lopkg_c(los, args_count_all_mono_class)
total_mono_all <- apply_count_to_lopkg_c(los, args_count_all_mono_all)
```

```{r}
tot_arg_df <- data.frame(
  Type=c("Full Monomorphic", "Monomorphic in Type", "Monomorphic in Class", "Monomorphic in Attribute Pattern", "Total Seen"),
  Count=c(total_mono_all, total_mono_type, total_mono_clas, total_mono_attr, total_arguments),
  Percentage=round(c(total_mono_all, total_mono_type, total_mono_clas, total_mono_attr, total_arguments) / total_arguments * 100, 2)
)

xtable(tot_arg_df)
```

```{r}
# function counts
total_functions <- apply_count_to_lopkg_c(los, length)
total_mono_type_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_type)
total_mono_attr_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_attr_pattern)
total_mono_clas_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_class)
total_mono_all_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_all)
```

```{r}
tot_fun_df <- data.frame(
  Type=c("Full Monomorphic", "Monomorphic in Type", "Monomorphic in Class", "Monomorphic in Attribute Pattern", "Total Seen"),
  Count=c(total_mono_all_functions, total_mono_type_functions, total_mono_clas_functions, total_mono_attr_functions, total_functions),
  Percentage=round(c(total_mono_all_functions, total_mono_type_functions, total_mono_clas_functions, total_mono_attr_functions, total_functions) / total_functions * 100, 2)
)

xtable(tot_fun_df)
```

We Have one thing we want to do first.
Namely, let's get rid of vector/X, scalar/X polymorphism.
```{r}
los_vs <- lapply(los, function(lofuns) {lapply(lofuns, combine_scalar_vector_where_appropriate)})
```

Top signatures, with printing.
This is how you get the top sigs.
```{r}
poly_type_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_type_args))
poly_type_args <- poly_type_args[order(-poly_type_args$count),] # sort by count
poly_type_args$perc <- round(poly_type_args$count/sum(poly_type_args$count) * 100, 2)
```

Print top 10.
```{r}
format_df_for_print(poly_type_args[1:10,]) -> put_me
xtable(put_me)
```

Now, having noted that:
(1) vector/X, scalar/X is common, and
(2) ints and doubles are basically interchangable
while noting that together these make up a huge portion of the polymorphism, we transform.
```{r}

```