### 2019-03-26 notebooks/corpus.Rmd
updated counts and new signatures, dunno if i should push the data (it big)

---
title: "Corpus Analysis"
output: html_notebook
---

Always load data processing.
```{r}
source("../code/data_processing.R")
```

```{r}
get_counts_for_pkg_in_counts_dir("rlang", path_to_counts = "../data/genthat_counts_final")
```

Get counts for observations.
```{r}

pkgs_to_count <- c("Rcpp", "rlang", "glue", "tibble", "stringi", "ggplot2", "dplyr", "pillar", "R6", "stringr")
names(pkgs_to_count) <- pkgs_to_count

top_obs <- lapply(pkgs_to_count, function(n) get_counts_for_pkg_in_counts_dir(n, path_to_counts = "../data/genthat_counts_final"))

top_obs

```

```{r}

```