### 2019-03-29 notebooks/corpus_stats_and_plots.Rmd
merged airplane changes. rip wow air

---
title: "Corpus Stats"
output: html_notebook
---

First order of business: the lines of code of the corpus.
Load.

```{r}
library(reshape2)
library(ggplot2)
library(ggrepel)
library(tidyverse)

count_df <- read_rds("../data/final_loc_counts.RDS")
# rows might need to be cast to integers
class(count_df$r_counts) <- "integer"
class(count_df$c_counts) <- "integer"

as.tibble(count_df) %>%
  mutate(names = rownames(count_df)) -> cloc

# for observations
t_r_s <- read_rds("../data/final_recorded_sigs_by_package.RDS")
tibble(names=names(t_r_s), rec = unlist(t_r_s)) -> total_recorded_sigs

r_b_f <- read_rds("../data/final_obs_counts_df.RDS")
as.tibble(r_b_f) %>%
  mutate(names = row.names(r_b_f)) %>%
  rename(obs = invo_obs) %>%
  select(names, obs) -> records_by_fun

total_recorded_sigs %>%
  inner_join(records_by_fun, by="names") %>%
  filter(obs != 0, rec !=0) %>% 
  inner_join(cloc, by="names") %>%
  filter(rec <= obs) %>%                      # Added this: FIXME, understand why
  filter(r_counts != 0) -> df

top_10 <- c("Rcpp", "rlang", "glue", "tibble", "stringi", "ggplot2", "dplyr", "pillar", "R6", "stringr")
# just named lists. some elements (in either) are integer(0) i believe, as some packages
# just have data.
```

First: let's get some of the counts for the beginning of the paper.
```{r}
tot_r <- sum(df$r_counts)
tot_c <- sum(df$c_counts)
nrow(df)
```

...I guess let's try to plot it??

```{r}
df %>%
  arrange(r_counts) %>%
  mutate(names_o = factor(names, names) , r_c_log= log10(r_counts), c_c_log = ifelse((c_counts==0),0, log10(c_counts)))  -> df_rc

df_rc %>% 
  ggplot(aes(x=names_o, y=r_c_log, group=1)) +
  geom_line() + geom_point() + labs(title = "Analysis of R Corpus (R Code Portion)",
    x = "Corpus of Analyzed Packages", 
    y = "Lines of R Code in Package (Log10-scaled)") + theme(axis.text.x = element_blank())
```

```{r}

df_rc %>% 
  ggplot(aes(x=names_o, y=c_c_log, group=1)) +
  geom_line() + geom_point() + labs(title = "Analysis of R Corpus (C Code Portion)",
    x = "Corpus of Analyzed Packages", 
    y = "Lines of C Code in Package (Log10-scaled)") + theme(axis.text.x = element_blank())
```

Similar to the R stuff.
Now, we should combine these plots.
Messy graphs, beware!
```{r}
#loc_of_top_10 <- lapply(1:length(combo_df$package_name), function(x) if (combo_df$package_name[[x]] %in% top_10) x else FALSE)

df_rc %>% 
  select(names_o, r_c_log, c_c_log) %>% 
  mutate(c_c_log=-c_c_log) %>%
  melt(id.vars="names_o",variable.name="language") -> zzz

zzz %>%
ggplot(aes(x=names_o, value, group=1)) +
 geom_line(aes(colour = language), size=0.008) + guides(colour=F) + 
 theme( axis.text.x  = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        axis.ticks.y = element_line(),  panel.grid.major.y = element_line(colour = "red")) + # ,
  scale_y_continuous(name="", breaks = c(-4, -3, -2, -1, 0, 1, 2, 3, 4), 
                     labels = c(10000, 1000, 100, 10, 0, 10, 100, 1000, 10000))

#+  geom_text_repel(data=subset(zzz, names %in% top_10 & language == "loc_r"), 
#                  aes(y = 0, label=names), position=position_dodge(0.9), direction="y")

ggsave("../paper/plots/linesofrandccode.pdf",width=10,height=4)
```

Plot:
```{r}
df_rc %>%
  mutate(rec=log10(rec),obs=log10(obs))  %>%
  ggplot(aes(y=rec, x=obs)) + geom_point(size=0.1) + 
    labs( x = "Observed", y = "Recorded") +
    scale_x_continuous(breaks = c( 1, 2, 3, 4, 5, 6, 7),
                       labels = c( 10, 100, 1000, 10000, 100000, 1000000, 10000000))+
    scale_y_continuous(breaks = c( 1, 2, 3, 4, 5, 6, 7),
                       labels = c( 10, 100, 1000, 10000, 100000, 1000000, 10000000))

# generate file
ggsave("../paper/plots/recordsbypkg.pdf",width=9,height=5)
```