### 2019-04-03 notebooks/jan.Rmd
minor

---
title: "Paper Results Section"
output: html_notebook
---

This Notebook walks through the results section of the paper.

First, administrivia.
```{r}
source("../code/data_processing.R") # get the functions we need

library(xtable)
```

Load the data.
```{r, echo=FALSE}
df <- list.files("../data/full_sigs", full.names=TRUE)

```