### 2019-04-03 notebooks/jan.Rmd
parser

---
title: "Paper Results Section"
output: html_notebook
---

This Notebook walks through the results section of the paper.

First, administrivia.
```{r}
source("../code/data_processing.R") # get the functions we need

library(xtable)
library(tidyverse)
```

Load the data.
```{r, echo=FALSE}
df <- readRDS("../data/per_sig_partial.RDS")
df <- as.tibble(df)
df %>% select(pkg, fun, arg1_t,  arg2_t, arg3_t, arg4_t, arg5_t, arg6_t, arg7_t,
              arg8_t, arg9_t, arg10_t, ret_t) %>% distinct() -> df2
df2
```