### 2019-04-03 notebooks/data_builder.Rmd
L1 translate code and merging

---
title: "Data Builder"
output: html_notebook
---

```{r}
source("../code/data_processing.R")
```

```{r}
# load the data
psp <- read_rds("../data/per_sig_partial.RDS")

# get rid of empties
# they are due to packages having no data, this may get resolved as the prl3 run finishes
psp <- psp[!psp$pkg == "no_data",]

# do this otherwise there are lists
psp <- lists_to_chars_sig_df(psp)
```

```{r}
# first, turn specials, closures, and builtins to functions
psp <- transform_new_df_w_TS_map(psp, type_map_to_fun)
```

```{r}
# make it T0
psp_L0 <- transform_new_df_w_TS_map(psp, type_map_T0_to_r)
```

```{r}
# write it to file for Java processing
write_csv(psp_L0, "../data/partial_L0.csv")
```

```{r}
# read in new one
# DONT FORGET TO GET RID OF }"" lol
psp_L0 <- read_csv("../data/partial_L0_subtype.csv")

# remove SUBTYPE rows
psp_L0 <- psp_L0[!psp_L0$pkg == "SUBTYPE",]
```

```{r}
# write L1 (which is just psp) to file for Java
write_csv(psp, "../data/partial_L1.csv")

# go to terminal, gzip it for Java
# DONT FORGET TO GET RID OF }"" lol
```

```{r}
# read it back
psp <- read_csv("../data/partial_L1_subtype.csv")

# remove SUBTYPE rows
psp <- psp[!psp$pkg == "SUBTYPE",]
```