### 2019-04-24 notebooks/rplus.Rmd
r plus, and updated java

---
title: "Tracing R Base: +"
output: html_notebook
---

Preamble.

```{r}
source("../code/data_processing.R")
```

In this notebook, you'll find our results from tracing part of the R base package.

Unfortunately, tracing base isn't going to be quite so straightforward.
The first and most relevant issue is that a number of the base package's functions are primitives (i.e. builtins) and are not amenable to the Genthat framework.
I needed to transform the + function into a closure in order to get this to shake out, and needed to make a number of light modifications to Genthat altogether.

For the record, this is what I needed to do to get + to trace:

```{r}
# unlock binding so we can write to it
# unlockBinding(as.symbol("+"), baseenv())
# This makes + into a closure, so we can deal with it
# assign("+", function(e1, e2) .Primitive("+")(e1, e2), envir=baseenv())
# Actually decorate the base env functions
# genthat::decorate_function(`+`, env=baseenv())
```

The above takes place in the genthat-tracing-site-file.

To load the data, run:

```{r}
# technically in L2
base_df <- readRDS("../data/some_plus_examples.RDS")

# for L0
base_df_L0 <- transform_new_df_w_TS_map(base_df, type_map_T0_to_r, unparam = T)

# just types
base_df_just_types <- base_df %>% select(arg1_t, arg2_t, ret_t)
```