### 2019-04-25 notebooks/refactor_test.Rmd
light modifications to some data files, new notebook, and new data

---
title: "Test Changes to Java Code"
output: html_notebook
---

This notebook is just to test the changes that were made to the Java code to make it more reasonable.
The main change was making Signatures contain a List of stuff, rather than just a bunch of fields.
It's easier to iterate through the fields this way, but a bit slower.

Ok. We ran Main with doIt uncommented.
What does L0_subtype look like?

```{r}
source("../code/data_processing.R")
```

```{r}
psp_L0 <- read_csv("../data/L0_subtype.csv.gz")
# ok that seems fine, try something else

psp_L0_col <- read_csv("../data/L0_collapsed.csv.gz")
# seems ok

L0_arg_sig_c <- read_csv("../data/L0_arg_sig_counts.csv.gz")
# seems fine

L0_sanitized <- read_csv("../data/L0_sanitized_class.csv.gz")
# yeah

psp_L2 <- read_csv("../data/L1_L2.csv.gz")
# yeah seems ok

list_of_L2_sig_c <- read_csv("../data/L1_arg_sig_counts_L2.csv.gz")
# yeah

```

Ok, things seem to have worked out.