### 2019-05-05 notebooks/rplus.Rmd
fixed optionals for brackets, and pushing data for that too

---
title: "Tracing R Base: +"
output: html_notebook
---

Preamble.

```{r}
source("../code/data_processing.R")
```

In this notebook, you'll find our results from tracing part of the R base package.

Unfortunately, tracing base isn't going to be quite so straightforward.
The first and most relevant issue is that a number of the base package's functions are primitives (i.e. builtins) and are not amenable to the Genthat framework.
I needed to transform the + function into a closure in order to get this to shake out, and needed to make a number of light modifications to Genthat altogether.

For the record, this is what I needed to do to get + to trace:

```{r}
# unlock binding so we can write to it
# unlockBinding(as.symbol("+"), baseenv())
# This makes + into a closure, so we can deal with it
# assign("+", function(e1, e2) .Primitive("+")(e1, e2), envir=baseenv())
# Actually decorate the base env functions
# genthat::decorate_function(`+`, env=baseenv())
```

The above takes place in the genthat-tracing-site-file.

To load the data, run:

```{r}
# technically in L2
# base_df <- readRDS("../data/some_plus_examples.RDS")
# base_df <- readRDS("../data/base_with_plus_and_brack.RDS")
base_df <- readRDS("../data/base_with_brack_fixed.RDS")

# for L0
base_df_L0 <- transform_new_df_w_TS_map(base_df, type_map_T0_to_r, unparam = T)
base_df_L0 <- unlist_new_df(base_df_L0)

# just types
base_df_just_types <- base_df %>% select(arg1_t, arg2_t, ret_t)
```

```{r}
# pass things over to java
# dont forget to tar

# save L0
write_csv(base_df_L0, "../data/plus_80_L0.csv")
# save L1
write_csv(base_df, "../data/plus_80_L1.csv")

# do the java

# read subtype
subtype_df_L0 <- read_csv("../data/plus_80_L0_subtype.csv.gz")
subtype_df_L0 <- subtype_df_L0[!subtype_df_L0$pkg == "SUBTYPE",]
subtype_df_L1 <- read_csv("../data/plus_80_L1_subtype.csv.gz")
subtype_df_L1 <- subtype_df_L1[!subtype_df_L1$pkg == "SUBTYPE",]
subtype_df_L2 <- read_csv("../data/plus_80_L1_L2.csv.gz")
subtype_df_L2 <- subtype_df_L2[!subtype_df_L2$pkg == "SUBTYPE",]
```

```{r}
# just get +

subtype_df_L0_p <- filter(subtype_df_L0, fun == "+") %>% salient_only_one_fun
subtype_df_L1_p <- filter(subtype_df_L1, fun == "+") %>% salient_only_one_fun
subtype_df_L2_p <- filter(subtype_df_L2, fun == "+") %>% salient_only_one_fun(just_types=F)

# just get [

subtype_df_L0_b <- filter(subtype_df_L0, fun == "[") %>% salient_only_one_fun(just_types = F)
subtype_df_L1_b <- filter(subtype_df_L1, fun == "[") %>% salient_only_one_fun(just_types = F)
subtype_df_L2_b <- filter(subtype_df_L2, fun == "[") %>% salient_only_one_fun(just_types=F)

```

] is somewhat interesting.
There were some issues with ].
The number of arguments that it takes is somewhat related to the dimensions of the thing being indexed.
For example, in data frames you can have a call to ] with 3 arguments: the data frame first, followed by the row and then column numbers (optionally).
For lists, it's always just 2 arguments.
Not sure how ... best to deal with this.
Right now, I just have it take 2 arguments: the thing being accessed, and an index variable.

There is perhaps room for a more sophisticateed analysis here, looking at the domensions of the first argument.
I don't think it'll be too challenging, but again, do we want to go that far?
I think that would be useful for some IDE tool or something: for instance, telling the programmer that they aren't using the thing with the right amount of dimensions.
Then again, you can totally do that and it's completely fine...

TODO: What package has the call with 1 argument?