### 2020-01-24 notebooks/data_showcase.Rmd
updated notebooks (by moving old ones into subdir), added new one, new code

---
title: "Data Showcase"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

The intent of this notebook is to showcase our current data analysis pipeline, and show some of the queries we indend to run.

```{r results='hide'}
source("../code/data_analysis_lib.R")
```

Load the data as `df`. `small_subset.csv` is around 50MB, so it should be manageable.

```{r results='hide'}
# This seems to be a fine size. We can run some easy stuff on it.
# The data is trimmed to only 20 arguments.
#
#
# 
# If the header isn't saved on the data:
# but you know how many arguments (or columns) are in the data, use
# either generate_header_for_num_args(num_args) or generate_col_names_for_count(num_columns)
df <- read_csv("../data/small_subset.csv")

# For some reason, the 
```

Now, here are some examples of the things that we might be interested in doing:

```{r}
# Get all `+` call traces:
#
# filter(fun_name == "+") returns all rows where the fun_name column is equal to "+".
#
# quick_trim_na_cols gets rid of columns entirely comprised of NAs. Useful when the function you're interested in
# doesn't have that many arguments, like `+`.
df %>% filter(fun_name == "+") %>% quick_trim_na_cols
```

Dealing with S3 dispatch:

```{r}
# How many function calls used the S3 dispatch mechanism?
df %>% filter(dispatch == "S3") %>% select(count) %>% sum -> num_S3

# You can also use some functions from the data_processing_lib imported at the top of this file.
# To get a breakdown of the dispatch usage in df, try:
df %>% breakdown_dispatch

# Takeaways:

# Breakdown by package:
df %>% breakdown_dispatch_by_pkg

# Takeaways:

# What classes were dispatched on?
# Note: Some of them might be {}, meaning no class. This is because whenever you call an S3 method, regardless of
#       the class of the first argument, it tries to dispatch. So we have cases where the class is empty ("{}")
#       which end up calling the generic (name.default) dispatch target.
df %>% breakdown_S3_dispatch_by_class

# Takeaways: most dispatch is done on data.frames, and POSIXct/POSIXt, then factors, then tibbles.
```

We also might be interested in instances of functions which have arguments which have multiple classes.

```{r}
# get_multi_classes adds columns of the form arg_cX_c, which has the # of classes in argument X of the function.
# 
df_multi <- df %>% get_multi_classes

# Breakdown by package::function:
df_multi %>% group_by(package, fun_name) %>% summarize
```

Will look into some of these and report results.