### 2020-05-06 evaluation/R/evaluate-type-analysis.Rmd
Add code for type coverage evaluation and type analysis and Rmd files

---
title: Evaluation of Types
author: Aviral Goel
output:
  html_document:
    theme: cerulean

params:
  type_analysis_datafile: ../data/type-analysis/merged/merged.csv
---

```{r setup, echo=FALSE, eval=TRUE}
#    toc: true
#    toc_float: true
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tibble))

type_analysis_data <- 
    read_csv(params$type_analysis_datafile,
             col_names = TRUE,
             col_types = cols(col_character(),
                              col_character(),
                              col_integer(),
                              col_integer(),
                              col_integer(),
                              col_character(),
                              col_integer()))

to_percentage <- function(proportions) {
    str_c(round(100 * proportions, 2), "%")
}
```

# High-level Summary

```{r high-level, echo=FALSE, eval=TRUE}
total_package_count <- nrow(distinct(type_analysis_data, package_name))
total_function_count <- nrow(distinct(type_analysis_data, package_name, function_name))
```

**Packages**: `r total_package_count`

**Functions**: `r total_function_count`

In a function declaration of the form:

`type f (<int | chr | lgl, dbl, lgl> => lgl | chr) | (<dbl | chr, int, raw> => lgl | chr)`

There are two outer alternatives and 3 inner alternatives for parameter 1.

# Outer Alternatives

Now, we look at the number of alternatives in the function type.

```{r outer_alternative_count, eval=TRUE, echo=FALSE}
outer_alternative_distribution <-
   type_analysis_data %>%
   group_by(package_name, function_name) %>%
   summarize(outer_alternative_count = length(unique(outer_alternative))) %>%
   ungroup() %>%
   group_by(outer_alternative_count) %>%
   summarize(function_count = n(), package_count = length(unique(package_name))) %>%
   ungroup()

datatable(outer_alternative_distribution)
```

# Inner Alternatives

## Parameters

Now, we look at the number of alternatives in the function argument type.

```{r argument_inner_alternative_count, eval=TRUE, echo=FALSE}
argument_inner_alternative_count <-
   type_analysis_data %>%
   dplyr::filter(parameter_position != -1L) %>%
   group_by(package_name, function_name, parameter_position) %>%
   summarize(inner_alternative_count = length(unique(inner_alternative))) %>%
   ungroup() %>%
   arrange(desc(inner_alternative_count))

datatable(argument_inner_alternative_count)
```

## Return

Now, we look at the number of alternatives in the function return type.

```{r result_inner_alternative_count, eval=TRUE, echo=FALSE}
result_inner_alternative_count <-
   type_analysis_data %>%
   dplyr::filter(parameter_position == -1L) %>%
   group_by(package_name, function_name) %>%
   summarize(inner_alternative_count = length(unique(inner_alternative))) %>%
   ungroup() %>%
   arrange(desc(inner_alternative_count))

result_inner_alternative_count %>%
dplyr::filter(inner_alternative_count != 1) %>%
datatable(caption="Number of union alternatives in return type")

result_inner_alternative_count_graph_data <-
result_inner_alternative_count %>%
group_by(inner_alternative_count) %>%
summarize(function_count = n()) %>%
ungroup()

result_inner_alternative_count %>%
mutate(inner_alternative_count = ceiling(inner_alternative_count / 5) * 5) %>%
group_by(inner_alternative_count) %>%
summarize(function_count = n(), package_count = length(unique(package_name))) %>%
ungroup() %>%
mutate(function_percentage = to_percentage(function_count / total_function_count),
       package_percentage = to_percentage(package_count / total_package_count)) %>%
arrange(desc(inner_alternative_count)) %>%
mutate(inner_alternative_count = str_c(inner_alternative_count - 4, 
                                       "-", 
                                       inner_alternative_count)) %>%
select(package_count, package_percentage, function_count, function_percentage, inner_alternative_count) %>%
datatable(caption="Distribution of return type alternatives")

explicit_cases <-
result_inner_alternative_count_graph_data %>%
filter(inner_alternative_count <= 5)

implicit_cases <-
result_inner_alternative_count_graph_data %>%
filter(inner_alternative_count > 5) %>%
summarize(function_count = length(function_count),
          inner_alternative_count = "> 5")

result_inner_alternative_count_graph_data <-
    explicit_cases %>%
    mutate(inner_alternative_count = as.character(inner_alternative_count)) %>%
    rbind(implicit_cases)

result_inner_alternative_count_graph_data %>%
ggplot(aes(inner_alternative_count, function_count)) +
geom_col() +
scale_x_discrete(limits = c(1:5, "> 5")) #%>%
#coord_flip()

```