### 2020-05-08 evaluation/R/monkey-patching.Rmd
Add analysis Rmds

---
title: "Monkey patching"
output: html_document
---

```{r setup, include=FALSE}
library(fs)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("paths.R")
```

## Fix parallel.log when running with `--resume-failed`

This will only keep the last run in the duplicated elements.

```{r}
parallel_log <- read_tsv(PACKAGE_COVERAGE_LOG)
count(parallel_log, Exitval)
updated_parallel_log <- group_by(parallel_log, Seq) %>% top_n(1, Starttime) %>% ungroup()
count(updated_parallel_log, Exitval)
write_tsv(updated_parallel_log, PACKAGE_COVERAGE_LOG)
```

## Fix package-coverage error code

```{r load coverage data}
df_coverage_run <- read_csv(PACKAGE_COVERAGE_RUN)
df_coverage <- read_csv(PACKAGE_COVERAGE)
```

```{r}
df_coverage_summary <- group_by(df_coverage, package) %>% summarise(success=sum(is.na(error)))

df_updated_coverage_run <- 
  df_coverage_run %>% 
  left_join(df_coverage_summary, by=c("job"="package")) %>% 
  mutate(exitval=case_when(
    exitval == 0 & success == 4 ~ 0,
    exitval == 0 & success != 4 ~ 1,
    TRUE ~ exitval
  )) %>%
  select(-success)
```

```{r}
count(df_updated_coverage_run, exitval)
```

```{r}
write_csv(df_updated_coverage_run, PACKAGE_COVERAGE_LOG)
```

```{r}
group_by(parallel_log, Seq) %>% top_n(1, -Starttime) %>% ungroup() %>% count(Exitval)
```

```{r}
parallel_log <- read_tsv(PACKAGE_COVERAGE_LOG)
updated_parallel_log <- left_join(parallel_log, select(df_coverage_run, Seq=seq, Exitval=exitval), by="Seq") %>%
  mutate(Exitval.x=Exitval.y) %>%
  rename(Exitval=Exitval.x) %>%
  select(-Exitval.y)
write_tsv(updated_parallel_log, PACKAGE_COVERAGE_LOG)
```