### 2020-05-08 evaluation/R/runtime-checks.Rmd

---
title: "Runtime checks"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(DT)
library(fs)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

```{r paths, include=FALSE}
source("inc/paths.R")
```

## Setup

The `r RUNTIME_CHECKS` file contains information about the use of assertions across packages.
To generate it, run the following:

```
 ./rapr/inst/on-each-package.sh ./rapr/inst/tasks/find-runtime-type-checks.R
```

or (with a `PACKAGES` variable set to restrict the package selection)

```
PACKAGES=packages-typer-35.txt ./rapr/inst/on-each-package.sh ./rapr/inst/tasks/find-runtime-type-checks.R
```

and then

```
env-rscript ./rapr/inst/process-csv-result.R run/find-runtime-type-checks runtime-checks.csv
```

to generate the CSV file.

## Data

```{r load run}
run <- read_csv(RUNTIME_CHECKS_RUN)
run <- semi_join(run, CORPUS_PACKAGES_DF, by=c("job"="package"))
```

```{r check run packages}
#stopifnot(nrow(run) == nrow(CORPUS_PACKAGES_DF))
```

```{r load runtime checks}
runtime_checks <- read_csv(RUNTIME_CHECKS)
runtime_checks <- semi_join(runtime_checks, CORPUS_PACKAGES_DF, by="package")
```

## Analysis

### By package

```{r}
per_package <- count(runtime_checks, package)
```

```{r}
per_package %>%
  arrange(desc(n)) %>%
  datatable()
```

```{r}
ggplot(per_package, aes(n)) + geom_histogram(binwidth = 1)
```

### By package function

```{r}
per_package_fun <- count(runtime_checks, package, fun_name)
```

```{r}
per_package_fun %>%
  arrange(desc(n)) %>%
  datatable()
```

```{r}
ggplot(per_package_fun, aes(n)) + geom_histogram(binwidth = 1)
```

### By asertion function

```{r}
per_assert <- count(runtime_checks, assert)
```

```{r}
per_assert %>%
  arrange(desc(n)) %>%
  datatable()
```

## Raw data

```{r}
datatable(
  runtime_checks, 
  filter="top", 
  options=list(
    pageLength=50,
    autoWidth=TRUE
  )
)
```