### 2020-05-08 evaluation/R/validation.Rmd

---
title: "Data and Validation"
output: html_notebook
---

Setup (this grabs Alexi's data analysis script. Ask him to move it if you're gonna use the notebook.)

```{r setup}
source("/mnt/home/pldi/scripts/support_scripts/data_analysis_lib.R")
```

Load the data.

```{r}
# df <- read_csv("/mnt/arraySSD/alexi/2020_03_14_candidate_reduced_20_simpl/all.csv")
df <- read_csv("/mnt/arraySSD/alexi/2020_03_14_candidate_reduced_20_simpl/all_twelfth.csv")
```

Now, onto the validation.
For this section, I'll include the queries that were used to validate the decisions we made when designing the type system.
The end goal will be to generate some latex to be included by the paper which should have the numbers, that way the paper is totally reproducible.
First, we get the data frame which counts instances of types.

```{r}
df_c <- df %>% quick_get_type_counts_with_row_nums
df_c <- df_c %>% ungroup
```

Let's start by breaking down the counts for tuples, lists, and structs.

```{r}
num_lists <- df_c %>% filter(substr(type, 1, 4) == "list") %>% select(count) %>% sum
# We don't have tuples per se, actually.
# num_tuples <- df_c %>% filter(substr(type, 1, 5) == "tuple") %>% select(count) %>% sum
# Data frames?
num_data_frame <- df_c %>% filter(substr(type, 1, 4) == "data") %>% select(count) %>% sum
num_structs <- df_c %>% filter(substr(type, 1, 6) == "struct") %>% select(count) %>% sum

total_listy <- num_lists + num_data_frame + num_structs

summary_listy <- data.frame(type = c("list", "data.frame", "struct"), count = c(num_lists, num_data_frame, num_structs), percentage = round(c(num_lists, num_data_frame, num_structs) / total_listy * 100, 2))

summary_listy
```

The above counts treat data.frames and structs differently, though they should be the same.
We also don't do tuples, they don't seem particularly relevant.

Let's grab counts for usage of scalar-variants of primitive types.

```{r}
df_c_just_primitives <- df_c %>% filter(substr(type, 1, 5) %in% substr(PRIM_TYPES, 1, 5))

num_scalars_no_names <- df_c_just_primitives %>% filter(type %in% paste0(PRIM_TYPES, "@NA-free")) %>% select(count) %>% sum
num_scalars_names <- df_c_just_primitives %>% from_counts_df_extract_dim_and_names %>% filter(num_names == 1) %>% select(count) %>% sum
num_vec_no_names <- df_c_just_primitives %>% filter(type %in% paste0(PRIM_TYPES, "[]") | type %in% paste0(PRIM_TYPES, "[]@NA-free")) %>% select(count) %>% sum
num_mat <- df_c_just_primitives %>% filter(type %in% paste0(PRIM_TYPES, "[,]") | type %in% paste0(PRIM_TYPES, "[,]@NA-free")) %>% select(count) %>% sum
num_vec_names <- df_c_just_primitives %>% from_counts_df_extract_dim_and_names %>% filter(num_names > 1) %>% select(count) %>% sum

total_prims <- df_c_just_primitives %>% select(count) %>% sum

summary_primitives <- data.frame(type = c("nameless scalar", "named scalar", "nameless vector", "named vector", "matrices"), count = c(num_scalars_no_names, num_scalars_names, num_vec_no_names, num_vec_names, num_mat), percentage = round( c(num_scalars_no_names, num_scalars_names, num_vec_no_names, num_vec_names, num_mat) / total_prims * 100, 2)) %>% as_tibble

summary_primitives
```

TODO: Look into why the percentages don't add up to 1. Maybe we missed something?

Next, let's try for ...

...

We can also synthesize type signatures from the data.
Let's try it:

First, an abstraction:

```{r}
get_fun_types_for_package <- function(df, pkg) {
  # First, filter out for a particular package.
  df_p <- df %>% filter(package == pkg)

  # What are the function names?
  p_fun_names <- df_p %>% select(fun_name) %>% unique %>% unlist %>% unname
  p_fun_names <- p_fun_names[!is.na(p_fun_names)]

  cat("p_fun_names: ")
  cat(p_fun_names, collapse=" ")
  cat("\n")

  # Let's build a type for each of these.
  names(p_fun_names) <- p_fun_names
  map(p_fun_names, function(x) {
    cat("processing ", x, "...\n");
    fn_id <- df_p %>% filter(fun_name == x) %>% select(fun_id) %>% unique %>% unlist %>% unname
    paste0(fn_id, ": ", get_type_of_function(df_p, pkg, x))
  })
}
```

Do it for some packages.
These are long-running, we split up the computation.

For dplyr:

```{r}
dplyr_types <- get_fun_types_for_package(df, "dplyr")
```

For the abind package:

```{r}
abind_types <- get_fun_types_for_package(df, "abind")
```

For the alluvial package:

```{r}
alluvial_types <- get_fun_types_for_package(df, "alluvial")
```

For the bayesplot package:

```{r}
bayesplot_types <- get_fun_types_for_package(df, "bayesplot")
```

```{r}
# Don't run this.
base_types <- get_fun_types_for_package(df, "base")
```