### 2020-05-15 paper/tables/uni-typed-breakdown.tex

\begin{table}[H]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lrrr}
\toprule
Types & Parameter \# & \% & Cumulative \%\\
\midrule
scalar & 28045 & 35.80 & 35.80\\
class & 18510 & 23.63 & 59.43\\
vector & 10296 & 13.14 & 72.58\\
\rowcolor{lightgray}  \textcolor{black}{...} & \textcolor{black}{6611} & \textcolor{black}{8.44} & \textcolor{black}{81.02}\\
\rowcolor{lightgray}  \textcolor{black}{any} & \textcolor{black}{5993} & \textcolor{black}{7.65} & \textcolor{black}{88.67}\\
\addlinespace
null & 4795 & 6.12 & 94.79\\
\rowcolor{lightgray}  \textcolor{black}{list} & \textcolor{black}{3229} & \textcolor{black}{4.12} & \textcolor{black}{98.91}\\
\textasciicircum{}vector & 461 & 0.59 & 99.50\\
environment & 390 & 0.50 & 100.00\\
\bottomrule
\end{tabular}}
\end{table}