### 2020-09-14 paper/tables/uni-typed-breakdown.tex
new words in. let the edits commence

\begin{table}[H]
  \centering
  \resizebox{.5\linewidth}{!}{
    \begin{tabular}{lrrr}
      \toprule
      Types & Parameter \# & \% & Cumulative \%\\
      \midrule
      scalar & 29673 & 35.06 & 35.06\\
      class & 20308 & 23.99 & 59.05\\
      vector & 11529 & 13.62 & 72.67\\
      \rowcolor{lightgray}  \textcolor{black}{...} & \textcolor{black}{6611} & \textcolor{black}{7.81} & \textcolor{black}{80.48}\\
      \rowcolor{lightgray} \textcolor{black}{null} & \textcolor{black}{6094} & \textcolor{black}{7.20} & \textcolor{black}{87.68}\\
        \addlinespace
   \rowcolor{lightgray}  \textcolor{black}{any} & \textcolor{black}{6011} & \textcolor{black}{7.10} & \textcolor{black}{94.78}\\
      \rowcolor{lightgray}  \textcolor{black}{list} & \textcolor{black}{3400} & \textcolor{black}{4.02} & \textcolor{black}{98.79}\\
      \textasciicircum{}vector & 612 & 0.72 & 99.52\\
      environment & 408 & 0.48 & 100.00\\
      \bottomrule
    \end{tabular}}
  \caption{Singleton Type Categories}\label{fig:singlefreq}
\end{table}