### 2020-09-14 paper/tables/uni-typed-breakdown.tex
major edits, minor edit pass now

\begin{table}[H]
  \centering
  \resizebox{.5\linewidth}{!}{
    \begin{tabular}{lrrr}
      \toprule
      Types & Parameter \# & \% & Cumulative \%\\
      \midrule
      scalar & 35064 & 33.33 & 33.33 \\
      class & 24256 & 23.06 & 56.39 \\
      vector & 13025 & 12.38 & 68.77 \\
      \rowcolor{lightgray}  \textcolor{black}{...} & \textcolor{black}{9142} & \textcolor{black}{8.69} & \textcolor{black}{77.46}\\
      \rowcolor{lightgray} \textcolor{black}{null} & \textcolor{black}{7694} & \textcolor{black}{7.31} & \textcolor{black}{84.77}\\
        \addlinespace
   \rowcolor{lightgray}  \textcolor{black}{any} & \textcolor{black}{7614} & \textcolor{black}{7.24} & \textcolor{black}{92.01}\\
      \rowcolor{lightgray}  \textcolor{black}{list} & \textcolor{black}{3558} & \textcolor{black}{3.38} & \textcolor{black}{95.39}\\
      \textasciicircum{}vector & 2923 & 2.78 & 98.17\\
      function & 1427 & 1.36 & 99.52 \\
      environment & 500 & 0.48 & 100.00\\
      \bottomrule
    \end{tabular}}
  \caption{Singleton Type Categories}\label{fig:singlefreq}
\end{table}